---
title: "Statistical Analysis of Active Layer Thickness in Utqiagvik, Alaska"
subtitle: "EES 4891/5891: Probability and Statistics for Geosciences"
author: "Maya Maciel-Seidman"
date: last-modified
---

## Setup

Set up R environment by loading the necessary libraries

```{r setup}
#| message: false
#| warning: false
#| echo: true
#| results: "hide"
library(knitr)
library(scales)
library(tidyverse)
library(ggplot2)
library(MASS, quietly = TRUE)

# maybe include other libraries:
# library(tidymodels)

# set up default plotting format for ggplot2 graphs
theme_set(theme_bw())
```

## Introduction

The data set that will be analyzed in this project is a data set of soil active layer thickness (the maximum depth of thawed soil above permafrost) and climate data for a study site in Utqiagvik, Alaska from 1995-2023. This data set was created last summer during a research internship I completed at the U.S. Naval Research Laboratory. This data is interesting to me because I am generally interested in the cryosphere and how climate change is affecting the Arctic, one of the most dynamic areas of the planet that is warming at rates 2-4 times faster than the rest of Earth.

### Data set

This data set has 29 observations of 11 variables. The measured variables include year, yearly thawing degree days (cumulative yearly temperature above 0 ℃), yearly freezing degree days (cumulative yearly temperature below  0 ℃), annual mean air temperature (measured in ℃), mean active layer thickness (measured in cm), annual mean snow depth (measured in in.), annual mean snow depth (measured in cm), annual total precipitation (only rain, measured in in,), annual total snowfall (measured in in.), and annual total snowfall (measured in cm). Only the metric measurements of variables will be used.

The active layer thickness measurements come from the Circumpolar Active Layer Monitoring (CALM) Network, an observational network that monitors the response of near-surface permafrost and soil active layer to climate change over multiple decades. The data for my specific study site in Utqiagvik is from CALM Site U1 Barrow, which is a 1 km x 1 km square grid of 121 sampling nodes. Active layer thickness is measured once a year at each of the 121 nodes and the yearly measurement in my data set is the average of these 121 measurements. The data for the climate variables in my data set were obtained from NOAA’s xmACIS, a data-query tool that provides access to climatological summary products from National Weather Service (NWS) field offices. It is directly linked to the Applied Climate Information System (ACIS) and Regional Climate Center (RCC) databases. I used xmACIS to obtain data from the NWS station located at Wiley Post-Will Rogers Memorial Airport in Utqiagvik, the closest station to the study site.

```{r}
alt_df <- read.csv("Copy of Utqiagvik Summary Dataset.csv")
alt_df <- alt_df %>% dplyr::select(!annual_mean_snow_depth_in)
alt_df <- alt_df %>% dplyr::select(!annual_total_precip_in)
alt_df <- alt_df %>% dplyr::select(!annual_total_snowfall_in)
alt_df <- alt_df %>% dplyr::select(!site_code)
alt_df <- alt_df %>% dplyr::select(!site_name)
```

## Descriptive Statistics

```{r}
# Yearly thawing degree days:
TDD_summary <- summary(alt_df$yearly_TDD)
TDD_summary["St. Dev"] <- sd(alt_df$yearly_TDD)
TDD_summary

ggplot(data= alt_df, aes(x = yearly_TDD)) +
  geom_density() +
  labs(x = "Yearly Thawing Degree Days", y = "Density", title = "Kernel Density of yearly_TDD")

ggplot(data = alt_df, aes(x = yearly_TDD)) +
  geom_histogram(colour = "black", fill = "darkgreen", bins = 20, binwidth = 30) +
  labs(x = "Yearly Thawing Degree Days", y = "Count", title = "Histogram of yearly_TDD")

ggplot(data = alt_df, aes(sample = yearly_TDD)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Normal Distribution Q-Q Plot for Yearly Thawing Degree Days")

TDD_gamma <- fitdistr(alt_df$yearly_TDD, densfun = "gamma")
ggplot(data = alt_df, aes(sample = yearly_TDD)) +
  geom_qq_line(distribution = qgamma, dparams = TDD_gamma$estimate) + 
  geom_qq(distribution = qgamma, dparams = TDD_gamma$estimate) +
  labs(title = "Gamma Distribution Q-Q Plot for Yearly Thawing Degree Days")

TDD_weibull <- fitdistr(alt_df$yearly_TDD, densfun = "weibull")
ggplot(data = alt_df, aes(sample = yearly_TDD)) +
  geom_qq_line(distribution = qweibull, dparams = TDD_weibull$estimate) + 
  geom_qq(distribution = qweibull, dparams = TDD_weibull$estimate) +
  labs(title = "Weibull Distribution Q-Q Plot for Yearly Thawing Degree Days")

TDD_poisson <- fitdistr(alt_df$yearly_TDD, densfun = "poisson")
ggplot(data = alt_df, aes(sample = yearly_TDD)) +
  geom_qq_line(distribution = qpois, dparams = TDD_poisson$estimate) + 
  geom_qq(distribution = qpois, dparams = TDD_poisson$estimate) +
  labs(title = "Poisson Distribution Q-Q Plot for Yearly Thawing Degree Days")
```

Based on the Q-Q plots, yearly thawing degree days does not seem to closely follow a normal, gamma, Weibull, or Poisson distribution. However, it follows a Weibull or Poisson distribution more closely than a normal or gamma distribution.

```{r}
# Yearly freezing degree days:
FDD_summary <- summary(alt_df$yearly_FDD)
FDD_summary["St. Dev"] <- sd(alt_df$yearly_FDD)
FDD_summary

ggplot(data= alt_df, aes(x = yearly_FDD)) +
  geom_density() +
  labs(x = "Yearly Freezing Degree Days", y = "Density", title = "Kernel Density of yearly_FDD")

ggplot(data = alt_df, aes(x = yearly_FDD)) +
  geom_histogram(colour = "black", fill = "darkgreen", bins = 20, binwidth = 30) +
  labs(x = "Yearly Freezinf Degree Days", y = "Count", title = "Histogram of yearly_FDD")

ggplot(data = alt_df, aes(sample = yearly_FDD)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Normal Distribution Q-Q Plot for Yearly Freezing Degree Days")
```

Based on the Q-Q plot, it seems that yearly freezing degree days does not follow a normal distribution. The other distributions could not be fit to this variable since all of the observations are negative numbers and the other distributions require non-negative numbers as inputs.

```{r}
# Annual mean air temperature:
air_temp_summary <- summary(alt_df$annual_mean_air_temp)
air_temp_summary["St. Dev"] <- sd(alt_df$annual_mean_air_temp)
air_temp_summary

ggplot(data= alt_df, aes(x = annual_mean_air_temp)) +
  geom_density() +
  labs(x = "Annual Mean Air Temperature", y = "Density", title = "Kernel Density of annual_mean_air_temp")

ggplot(data = alt_df, aes(x = annual_mean_air_temp)) +
  geom_histogram(colour = "black", fill = "darkgreen", bins = 30, binwidth = 1) +
  labs(x = "Annual Mean Air Temperature", y = "Count", title = "Histogram of annual_mean_air_temp")

ggplot(data = alt_df, aes(sample = annual_mean_air_temp)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Normal Distribution Q-Q Plot for Annual Mean Air Temperature")
```

Looking at the Q-Q plot, annual mean air temperature does not particularly follow a normal distribution. The other distributions could not be fit to this variable since all of the observations are negative numbers and the other distributions require non-negative numbers as inputs.

```{r}
# Active layer thickness:
ALT_summary <- summary(alt_df$mean_ALT)
ALT_summary["St. Dev"] <- sd(alt_df$mean_ALT)
ALT_summary

ggplot(data= alt_df, aes(x = mean_ALT)) +
  geom_density() +
  labs(x = "Active Layer Thickness", y = "Density", title = "Kernel Density of mean_ALT")

ggplot(data = alt_df, aes(x = mean_ALT)) +
  geom_histogram(colour = "black", fill = "darkgreen", bins = 20, binwidth = 1.75) +
  labs(x = "Active Layer Thickness", y = "Count", title = "Histogram of mean_ALT")

ggplot(data = alt_df, aes(sample = mean_ALT)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Normal Distribution Q-Q Plot for Active Layer Thickness")

ALT_gamma <- fitdistr(alt_df$mean_ALT, densfun = "gamma")
ggplot(data = alt_df, aes(sample = mean_ALT)) +
  geom_qq_line(distribution = qgamma, dparams = ALT_gamma$estimate) + 
  geom_qq(distribution = qgamma, dparams = ALT_gamma$estimate) +
  labs(title = "Gamma Distribution Q-Q Plot for Active Layer Thickness")

ALT_weibull <- fitdistr(alt_df$mean_ALT, densfun = "weibull")
ggplot(data = alt_df, aes(sample = mean_ALT)) +
  geom_qq_line(distribution = qweibull, dparams = ALT_weibull$estimate) + 
  geom_qq(distribution = qweibull, dparams = ALT_weibull$estimate) +
  labs(title = "Weibull Distribution Q-Q Plot for Active Layer Thickness")

ALT_poisson <- fitdistr(alt_df$mean_ALT, densfun = "poisson")
ggplot(data = alt_df, aes(sample = mean_ALT)) +
  geom_qq_line(distribution = qpois, dparams = ALT_poisson$estimate) + 
  geom_qq(distribution = qpois, dparams = ALT_poisson$estimate) +
  labs(title = "Poisson Distribution Q-Q Plot for Active Layer Thickness")
```

Based on the Q-Q plots, it appears that active layer thickness most closely follows a gamma distribution.

```{r}
# Snow depth:
snow_depth_summary <- summary(alt_df$annual_mean_snow_depth_cm)
snow_depth_summary["St. Dev"] <- sd(alt_df$annual_mean_snow_depth_cm)
snow_depth_summary

ggplot(data= alt_df, aes(x = annual_mean_snow_depth_cm)) +
  geom_density() +
  labs(x = "Snow Depth", y = "Density", title = "Kernel Density of annual_mean_snow_depth_cm")

ggplot(data = alt_df, aes(x = yearly_TDD)) +
  geom_histogram(colour = "black", fill = "darkgreen", bins = 20, binwidth = 30) +
  labs(x = "Yearly Thawing Degree Days", y = "Count", title = "Histogram of yearly_TDD")

ggplot(data = alt_df, aes(sample = annual_mean_snow_depth_cm)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Normal Distribution Q-Q Plot for Snow Depth")
```

Annual mean snow depth does not closely follow a normal distribution, based on the Q-Q plot. I couldn't fit the other distributions to the data since there exists missing data. I chose not to omit the missing data since my data set does not have many observations and I did not want to decrease the number of observations and make the data set even smaller.

```{r}
# Precipitation:
precip_summary <- summary(alt_df$annual_total_precip_cm)
precip_summary["St. Dev"] <- sd(alt_df$annual_total_precip_cm)
precip_summary

ggplot(data= alt_df, aes(x = annual_total_precip_cm)) +
  geom_density() +
  labs(x = "Annual Total Precipitation", y = "Density", title = "Kernel Density of annual_total_precip_cm")

ggplot(data = alt_df, aes(x = annual_total_precip_cm)) +
  geom_histogram(colour = "black", fill = "darkgreen", bins = 20, binwidth = 2) +
  labs(x = "Annual Total Precipitation", y = "Count", title = "Histogram of annual_total_precip_cm")

ggplot(data = alt_df, aes(sample = annual_total_precip_cm)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Normal Distribution Q-Q Plot for Annual Total Precipitation")

precip_gamma <- fitdistr(alt_df$annual_total_precip_cm, densfun = "gamma")
ggplot(data = alt_df, aes(sample = annual_total_precip_cm)) +
  geom_qq_line(distribution = qgamma, dparams = precip_gamma$estimate) + 
  geom_qq(distribution = qgamma, dparams = precip_gamma$estimate) +
  labs(title = "Gamma Distribution Q-Q Plot for Annual Total Precipitation")

precip_weibull <- fitdistr(alt_df$annual_total_precip_cm, densfun = "weibull")
ggplot(data = alt_df, aes(sample = annual_total_precip_cm)) +
  geom_qq_line(distribution = qweibull, dparams = precip_weibull$estimate) + 
  geom_qq(distribution = qweibull, dparams = precip_weibull$estimate) +
  labs(title = "Weibull Distribution Q-Q Plot for Annual Total Precipitation")

precip_poisson <- fitdistr(alt_df$annual_total_precip_cm, densfun = "poisson")
ggplot(data = alt_df, aes(sample = annual_total_precip_cm)) +
  geom_qq_line(distribution = qpois, dparams = precip_poisson$estimate) + 
  geom_qq(distribution = qpois, dparams = precip_poisson$estimate) +
  labs(title = "Poisson Distribution Q-Q Plot for Annual Total Precipitation")
```

Based on the Q-Q plots, a gamma distribution seems to be the best fit for annual total precipitation.

```{r}
# Snowfall:
snowfall_summary <- summary(alt_df$annual_total_snowfall_cm)
snowfall_summary["St. Dev"] <- sd(alt_df$annual_total_snowfall_cm)
snowfall_summary

ggplot(data= alt_df, aes(x = annual_total_snowfall_cm)) +
  geom_density() +
  labs(x = "Annual Total Snowfall", y = "Density", title = "Kernel Density of annual_total_snowfall_cm")

ggplot(data = alt_df, aes(x = annual_total_snowfall_cm)) +
  geom_histogram(colour = "black", fill = "darkgreen", bins = 20, binwidth = 20) +
  labs(x = "Annual Total Snowfall", y = "Count", title = "Histogram of annual_total_snowfall_cm")

ggplot(data = alt_df, aes(sample = annual_total_snowfall_cm)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Normal Distribution Q-Q Plot for Annual Total Snowfall")
```

The Q-Q plot conveys that annual total snowfall does not closely follow a normal distribution. I again chose not to remove missing data, so I did not fit the other distributions.

## Research Questions

1) How are individual climate variables correlated with active layer thickness?

2) Which climate variable is the most strongly correlated with active layer thickness?

3) How well does linear regression work to predict active layer thickness, using each individual climate variable?

4) Which climate variable is the best to use in a linear regression model to predict active layer thickness?

## Hypotheses and Tests

### Hypotheses

1) Yearly thawing degree days, annual mean air temperature, and annual total precipitation will be positively correlated with active layer thickness. Yearly freezing degree days, annual mean snow depth, and annual total snowfall will be negatively correlated with active layer thickness.

2) Annual mean air temperature will have the strongest correlation with active layer thickness.

3) Linear regression will work well for yearly thawing degree days, annual mean air temperature, yearly freezing degree days, and annual total snowfall because I hypothesize that these variables will have a linear relationship with active layer thickness.

4) Annual mean air temperature will be the best variable for linear regression because I hypothesize that the relationship between annual mean air temperature and active layer thickness will be the most linear out of the climate variables.

### Methods

To analyze the correlation between each individual climate variable and active layer thickness, I will be using Spearman correlations because the data is not all normally distributed, which means that a Pearson correlation may not be the best choice since it assumes that both variables are normally distributed. On the other hand, data does not need to be normally distributed for Spearman correlations.

For the linear regressions, I will be using `lm()` to create a linear model with each climate variable and active layer thickness. I will use the models to make predictions and compared the predicted values of active layer thickness to the observed values of active layer thickness. I will also analyze the normality and homogeneity of each model's residuals.

### Hypothesis Tests

#### Spearman Correlations:

```{r}
# Yearly thawing degree days:
cor.test(alt_df$yearly_TDD, alt_df$mean_ALT, method = "spearman")
```

Since rho is 0.65, yearly thawing degree days has a strong positive relationship with active layer thickness. Since the p-value is less than 0.05, this relationship is statistically significant.

```{r}
# Yearly freezing degree days:
cor.test(alt_df$yearly_FDD, alt_df$mean_ALT, method = "spearman")
```

Since rho is 0.35, yearly freezing degree days has a moderate positive relationship with active layer thickness. The p-value is greater than 0.05, though, so this relationship is not statistically significant.

```{r}
# Annual mean air temperature:
cor.test(alt_df$annual_mean_air_temp, alt_df$mean_ALT, method = "spearman")
```

With a rho of 0.46, annual mean air temperature has a strong positive relationship with active layer thickness. The p-value is less than 0.05, indicating that this relationship is statistically significant.

```{r}
# Annual mean snow depth:
cor.test(alt_df$annual_mean_snow_depth_cm, alt_df$mean_ALT, method = "spearman")
```

The rho is -0.417, so the relationship between annual mean snow depth and active layer thickness is strong and negative. The p-value is less than 0.05, so this relationship is statistically significant.

```{r}
# Annual total precipitation:
cor.test(alt_df$annual_total_precip_cm, alt_df$mean_ALT, method = "spearman")
```

The rho of 0.42 indicates a strong positive relationship between annual total precipitation and active layer thickness. The p-value is less than 0.05, which indicates that this relationship is statistically significant.

```{r}
# Annual total snowfall:
cor.test(alt_df$annual_total_snowfall_cm, alt_df$mean_ALT, method = "spearman")
```

With a rho of -0.28, there is a weak negative relationship between annual total snowfall and active layer thickness. The p-value is greater than 0.05, so this relationship is not statistically significant.

#### Linear Regressions:

##### Yearly Thawing Degree Days

```{r}
tdd_model <- lm(mean_ALT ~ yearly_TDD, data = alt_df)

summary(tdd_model)
```

TODO: interpret this

```{r}
alt_df <- alt_df %>% mutate(tdd_pred = predict(tdd_model, newdata = alt_df))

ggplot(alt_df, aes(x = yearly_TDD, y = tdd_pred)) +
  geom_point(color = "darkblue") +
  geom_point(aes(y = mean_ALT), color = "darkred") +
  labs(x = "Yearly Thawing Degree Days", y = "Active Layer Thickness", title = "Predicted vs. Observed ALT Based on yearly_TDD")
```

TODO: interpret this

```{r}
tdd_res <- alt_df %>% filter(! is.na(mean_ALT)) %>% 
  mutate(res = mean_ALT - tdd_pred)

ggplot(tdd_res, aes(sample = res)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Q-Q Plot for TDD Residuals")
```

TODO: interpret this

```{r}
ggplot(tdd_res, aes(x = mean_ALT, y = yearly_TDD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Active Layer Thickness", y = "Yearly Thawing Degree Days", title = "Modeling ALT from Yearly TDD")

ggplot(tdd_res, aes(x = mean_ALT, y = res)) +
  geom_point() +
  labs(x = "Active Layer Thickness", y = expression(epsilon), title = "Residuals")
```

TODO: interpret this

##### Yearly Freezing Degree Days

```{r}
fdd_model <- lm(mean_ALT ~ yearly_FDD, data = alt_df)

summary(fdd_model)
```

TODO: interpret this

```{r}
alt_df <- alt_df %>% mutate(fdd_pred = predict(fdd_model, newdata = alt_df))

ggplot(alt_df, aes(x = yearly_FDD, y = fdd_pred)) +
  geom_point(color = "darkblue") +
  geom_point(aes(y = mean_ALT), color = "darkred") +
  labs(x = "Yearly Freezing Degree Days", y = "Active Layer Thickness", title = "Predicted vs. Observed ALT Based on yearly_FDD")
```

TODO: interpret this

```{r}
fdd_res <- alt_df %>% filter(! is.na(mean_ALT)) %>% 
  mutate(res = mean_ALT - fdd_pred)

ggplot(fdd_res, aes(sample = res)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Q-Q Plot for FDD Residuals")
```

TODO: interpret this

```{r}
ggplot(fdd_res, aes(x = mean_ALT, y = yearly_FDD)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Active Layer Thickness", y = "Yearly Freezing Degree Days", title = "Modeling ALT from Yearly FDD")

ggplot(fdd_res, aes(x = mean_ALT, y = res)) +
  geom_point() +
  labs(x = "Active Layer Thickness", y = expression(epsilon), title = "Residuals")
```

TODO: interpret this

##### Annual Mean Air Temperature

```{r}
air_temp_model <- lm(mean_ALT ~ annual_mean_air_temp, data = alt_df)

summary(air_temp_model)
```

TODO: interpret this

```{r}
alt_df <- alt_df %>% mutate(air_temp_pred = predict(air_temp_model, newdata = alt_df))

ggplot(alt_df, aes(x = annual_mean_air_temp, y = air_temp_pred)) +
  geom_point(color = "darkblue") +
  geom_point(aes(y = mean_ALT), color = "darkred") +
  labs(x = "Annual Mean Air Temperature", y = "Active Layer Thickness", title = "Predicted vs. Observed ALT Based on annual_mean_air_temp")
```

TODO: interpret this

```{r}
air_temp_res <- alt_df %>% filter(! is.na(mean_ALT)) %>% 
  mutate(res = mean_ALT - air_temp_pred)

ggplot(air_temp_res, aes(sample = res)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Q-Q Plot for Air Temperature Residuals")
```

TODO: interpret this

```{r}
ggplot(air_temp_res, aes(x = mean_ALT, y = annual_mean_air_temp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Active Layer Thickness", y = "Annual Mean Air Temperature", title = "Modeling ALT from Annual Mean Air Temperature")

ggplot(air_temp_res, aes(x = mean_ALT, y = res)) +
  geom_point() +
  labs(x = "Active Layer Thickness", y = expression(epsilon), title = "Residuals")
```

TODO: interpret this

##### Annual Mean Snow Depth

```{r}
snow_depth_model <- lm(mean_ALT ~ annual_mean_snow_depth_cm, data = alt_df)

summary(snow_depth_model)
```

TODO: interpret this

```{r}
alt_df <- alt_df %>% mutate(snow_depth_pred = predict(snow_depth_model, newdata = alt_df))

ggplot(alt_df, aes(x = annual_mean_snow_depth_cm, y = snow_depth_pred)) +
  geom_point(color = "darkblue") +
  geom_point(aes(y = mean_ALT), color = "darkred") +
  labs(x = "Annual Mean Snow Depth", y = "Active Layer Thickness", title = "Predicted vs. Observed ALT Based on annual_mean_snow_depth_cm")
```

TODO: interpret this

```{r}
snow_depth_res <- alt_df %>% filter(! is.na(mean_ALT)) %>% 
  mutate(res = mean_ALT - snow_depth_pred)

ggplot(snow_depth_res, aes(sample = res)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Q-Q Plot for Snow Depth Residuals")
```

TODO: interpret this

```{r}
ggplot(snow_depth_res, aes(x = mean_ALT, y = annual_mean_snow_depth_cm)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Active Layer Thickness", y = "Annual Mean Snow Depth", title = "Modeling ALT from Annual Mean Snow Depth")

ggplot(snow_depth_res, aes(x = mean_ALT, y = res)) +
  geom_point() +
  labs(x = "Active Layer Thickness", y = expression(epsilon), title = "Residuals")
```

TODO: interpret this

##### Annual Total Precipitation

```{r}
precip_model <- lm(mean_ALT ~ annual_total_precip_cm, data = alt_df)

summary(precip_model)
```

TODO: interpret this

```{r}
alt_df <- alt_df %>% mutate(precip_pred = predict(precip_model, newdata = alt_df))

ggplot(alt_df, aes(x = annual_total_precip_cm, y = precip_pred)) +
  geom_point(color = "darkblue") +
  geom_point(aes(y = mean_ALT), color = "darkred") +
  labs(x = "Annual Total Precipitation", y = "Active Layer Thickness", title = "Predicted vs. Observed ALT Based on annual_total_precip_cm")
```

TODO: interpret this

```{r}
precip_res <- alt_df %>% filter(! is.na(mean_ALT)) %>% 
  mutate(res = mean_ALT - precip_pred)

ggplot(precip_res, aes(sample = res)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Q-Q Plot for Precipitation Residuals")
```

TODO: interpret this

```{r}
ggplot(precip_res, aes(x = mean_ALT, y = annual_total_precip_cm)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Active Layer Thickness", y = "Annual Total Precipitation", title = "Modeling ALT from Yearly TDD")

ggplot(precip_res, aes(x = mean_ALT, y = res)) +
  geom_point() +
  labs(x = "Active Layer Thickness", y = expression(epsilon), title = "Residuals")
```

TODO: interpret this

##### Annual Total Snowfall

```{r}
snowfall_model <- lm(mean_ALT ~ annual_total_snowfall_cm, data = alt_df)

summary(snowfall_model)
```

TODO: interpret this

```{r}
alt_df <- alt_df %>% mutate(snowfall_pred = predict(snowfall_model, newdata = alt_df))

ggplot(alt_df, aes(x = annual_total_snowfall_cm, y = snowfall_pred)) +
  geom_point(color = "darkblue") +
  geom_point(aes(y = mean_ALT), color = "darkred") +
  labs(x = "Annual Total Snowfall", y = "Active Layer Thickness", title = "Predicted vs. Observed ALT Based on annual_total_snowfall_cm")
```

TODO: interpret this

```{r}
snowfall_res <- alt_df %>% filter(! is.na(mean_ALT)) %>% 
  mutate(res = mean_ALT - snowfall_pred)

ggplot(snowfall_res, aes(sample = res)) +
  geom_qq_line() + geom_qq() +
  labs(title = "Q-Q Plot for Snowfall Residuals")
```

TODO: interpret this

```{r}
ggplot(snowfall_res, aes(x = mean_ALT, y = annual_total_snowfall_cm)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Active Layer Thickness", y = "Annual Total Snowfall", title = "Modeling ALT from Annual Total Snowfall")

ggplot(snowfall_res, aes(x = mean_ALT, y = res)) +
  geom_point() +
  labs(x = "Active Layer Thickness", y = expression(epsilon), title = "Residuals")
```

TODO: interpret this

### Results and Discussion

Results from your hypothesis tests and discussion, interpreting your results

#### Spearman Correlations

TODO: which ones are positive and negative, which ones are significant and not, surprising results

#### Linear Regression Models

TODO: when linear regression worked well and when it didn't, most promising variable for linear regression

## Conclusions

What did you learn? What results would a reader be interested in? What would be the next steps if you were to continue studying this data after the end of the semester?
